---
title: "Rianne Schouten"
subtitle: "PhD Candidate"
author: |
  | Exceptional Model Mining | Missing Data | Causal Inference
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

---

### Introduction and affiliations

My name is Rianne and I analyze analysis methods. 

I am a PhD Candidate at Eindhoven University of Technology and study the Exceptional Model Mining framework. We work on developing solutions for heterogeneous datasets with time-dependent attributes. My supervisors are [dr. Wouter Duivesteijn](http://wwwis.win.tue.nl/~wouter/) and [prof. Mykola Pechenizkiy](https://www.win.tue.nl/~mpechen/). 

Additionally, I work on Missing Data and Causal Inference methods. Under supervision of [prof. Stef van Buuren](https://scholar.google.nl/citations?user=_3y5C0UAAAAJ&hl=nl) and [dr. Gerko Vink](https://www.gerkovink.com/), I co-developed a procedure to generate missing data in complete datasets and implemented the method in R-function [`ampute`](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html). It is part of R-package [`mice`](https://github.com/stefvanbuuren/mice).

I find it important that my work can be used for real-world problems, especially those in the health care domain. I furthermore enjoy collaborating on projects where data analysis and pattern mining meet practical problems. Please contact me if you have questions or seek opportunities to work together. 

---

### Publications

- Van der Haar, J.F., Nagelkerken, S.C., Smit, I.G., van Straaten, K., Tack, J.A., Schouten, R.M. & Duivesteijn, W. (2022) [Efficient Subgroup Discovery Through Auto-Encoding.](pdfs/2022_vdHaar_AE.pdf) In: Proceedings of the 20th International Symposium on Intelligent Data Analysis (IDA), pp. 327-340.

- Schouten, R.M., Duivesteijn, W. & Pechenizkiy, M. (2022) [Exceptional Model Mining for Repeated Cross-Sectional Data (EMM-RCS)](pdfs/2022_Schouten_EMMRCS.pdf). In: Proceedings of SIAM International Conference on Data Mining (SDM), pp. 585-593.

- Schouten, R.M., Bueno, M.L.P., Duivesteijn, W. & Pechenizkiy, M. (2022) [Mining Sequences with Exceptional Transition Behaviour of Varying Order using Quality Measures based on Information-Theoretic Scoring Functions.](pdfs/MiningExceptionalSequences_DAMI_2022.pdf) Data Mining and Knowledge Discovery (DAMI), 36: 379-413.

- IJsselhof R, Duchateau S, Schouten R, Slieker M, Hazekamp M & Schoof P. [Long-Term Follow-Up of Pericardium
for the Ventricular Component in Atrioventricular Septal Defect Repair.](https://pubmed.ncbi.nlm.nih.gov/33164688/) World Journal for Pediatric and Congenital Heart Surgery, 11(6): 742-747.

- IJsselhof R.J., Duchateau S.D.R., Schouten R.M., Freund, M.W., Heuser, J., Fejzic, Z., Haas, F., Schoof, P.H. & Slieker, M.G. (2019) [Follow-up After Biventricular Repair of the Hypoplastic Left Heart Complex.](https://academic.oup.com/ejcts/article-abstract/57/4/644/5607276) European Journal of Cardiothoracic Surgery, 57(4): 644-651.

- Rianne R.M., Lugtig, P. & Vink, G. (2018) [Generating missing values for simulation purposes: A multivariate amputation procedure.](https://www.tandfonline.com/doi/full/10.1080/00949655.2018.1491577) Journal of Statistical Computation and Simulation, 88(15): 1909-1930.

- Schouten, R.M. and Vink, G. (2021) [The dance of the mechanisms: How observed information influences the validity of missingness assumptions.](https://journals.sagepub.com/doi/full/10.1177/0049124118799376) Sociological Methods \& Research, 50(3): 1243-1258.

- Kappen, I.F.P.M., Bittermann, G.K.P., Schouten, R.M., Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. & Breugem, C.C. (2017) [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis.](https://www.researchgate.net/publication/308181181_Long-term_mid-facial_growth_of_patients_with_a_unilateral_complete_cleft_of_lip_alveolus_and_palate_treated_by_two-stage_palatoplasty_cephalometric_analysis) Clinical Oral Investigations, 21: 1801-1810.

- de Vries, C.P., Schouten, R.M., Van der Kuur, J., Gottardi, L., & Akamatsu, H. (2016) [Microcalorimeter pulse analysis by means of principle component decomposition.](https://arxiv.org/abs/1607.00834) Proceedings SPIE 9905, Space TElescopes and Instrumentation 2016: Ultraviolet to Gamma Ray, 99055v. DOI: 10.1117/12.2231627

---

### Software development

#### 1. `ampute` in R-package **mice**

```{r, eval = FALSE}
library(mice)
?ampute
```

R-function `ampute` is the implementation of a multivariate amputation procedure: a method for generating missing data in complete datasets. With `ampute`, it is straightforward to generate missing values in multiple variables, with different missing data proportions and varying underlying missingness mechanisms. Read [the article](pdfs/ArticleAmpute.pdf) or [the vignette](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html) to learn more.

#### 2. `parlMICE` 

For large datasets or when you want to impute with a large number of imputations, multiple imputation with `mice` in R-package **mice** may have a long run time. As a solution, Gerko Vink and I created wrapper function `parlMICE`, which allows for a parallel run of `mice`.

The function is now part of package **mice** under the name `parlmice`.

```{r, eval = FALSE}
library(mice)
?parlmice
```

All information can be found [in the github repo](https://github.com/RianneSchouten/parlMICE) or in the [vignette](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html).

#### 3. `pyampute`: the first Python library for data amputation

Library `pyampute` provides the multivariate amputation methodology to the Python community, and it does more. It has improved default settings, allows for a combined MAR+MNAR mechanism, for custom probability functions and since it is compatible with scikit-learn's `fit` and `transform` paradigm, seamless integration in data processing pipelines becomes easy. 

Find plenty of examples in [`pyampute`'s documentation](https://rianneschouten.github.io/pyampute/build/html/index.html). Install using pip or from source:

```
pip install pyampute
```

```
git clone https://github.com/RianneSchouten/pyampute.git
pip install ./pyampute
```

---

### Invited presentations

##### * EAISI Eindhoven, November 2021

##### * Workshop R-Ladies Amsterdam, October 2019

[Find the materials here](https://github.com/RianneSchouten/workshops/tree/master/R-Ladies%20Workshop)

The workshop covers: 

1.	Missing Values Analysis
2.	Evaluating Missing Data methods
3.	Implementing Missing Data Methods

##### * Data Science Hackathon, January 2018

My presentation can be found [here](pdfs/Data_Science_Hackathon_Rianne_Schouten.pdf).

##### * Amst-R-Dam 2017

A presentation about amputation and how to use `ampute` to do this. 

### Contributed presentations

##### * Poster at ECML-PKDD 2021 Satellite Event

Find the poster [here](pdfs/PosterSatelliteEvent_RMSchoutenetal.pdf).

##### * The Neglected Assumptions in Causal Inference (NACI) workshop at International Conference on Machine Learning (ICML), July 2021

Find the poster [here](pdfs/NACI_Poster_RMSchouten.pdf).

##### * Masterclass @ European Women in Technology, November 2018

Dealing with missing data in R: Amputation or Imputation? Find here the [presentation and exercises](https://github.com/RianneSchouten/workshops/tree/master/WomenInTechnology).

##### * sat-R-Day, September 2018

An explanation of missing data mechanisms, missing data patterns and how to 
generate them with `ampute`. Check the [slides](pdfs/SatRdayRianneSchouten.pdf)

##### * ICT Open, March 2018

Handling Missing Data in Data Science: Simulating the effects of missing data methods and how to present the results in an interactive plot with Github Pages. Read the [abstract](pdfs/abstract_rianne_schouten.pdf) or check the [slides](pdfs/ictopen2018_RianneSchouten.pdf)

##### * useR!2017 Brussels, July 2017

My presentation discussed the generation of missing values with `ampute`. A [summary](https://user2017.sched.com/event/AxsS?iframe=no) can be found on the website of the conference. 

---

### Contact

- Email: r.m.schouten\@tue.nl

- Github: [/RianneSchouten](https://github.com/RianneSchouten)

- [LinkedIn](https://www.linkedin.com/in/rianne-schouten-5bb29491/)

- Twitter: [missD_ta](https://twitter.com/missD_ta)

- [Google Scholar](https://scholar.google.nl/citations?view_op=list_works&hl=nl&user=xC56bhMAAAAJ&gmla=AJsN-F4IWn_9Vw9hH7hPN1zIMfNadBL6mvmu-rPpRS6p-4Enc7fR6GrOMYyN2QEr21D5Y3Uhjtj0HPT3I7qsh3MVE1v2L8MRfjsk7W1Qop4mDunpnTRDeMYF0DmoRYmja6b7GvnmC4R3)

- [Research Gate](https://www.researchgate.net/profile/Rianne_Schouten/publications)

- [Curriculum Vitae](pdfs/Rianne_Schouten_Resume.pdf)

