---
title: "Rianne Schouten"
author: |
  | PhD Candidate / Missing Data Specialist / Junior Data Scientist
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

---

### Introduction and affiliations

I am an enthusiastic person, eager to learn, 

I am doing research under supervision of [Gerko Vink](www.gerkovink.com) and [Stef van Buuren](https://github.com/stefvanbuuren) at Department of Methodology and Statistics, University of Utrecht. A research [proposal] further introduces the contents of my research.

I work as a missing data specialist and junior data scientist at DPA Professionals. I am part of the [Data Science Excellence Program](www.dpadatascience.nl), lead by [Fred Kroon](https://www.linkedin.com/in/kroonfred/). We are trained by senior data scientists from BDR. DPA funds my research by giving me time to do my research. 

---

### Blogposts

- [Inzicht in je missing data](https://www.dpa.nl/it/blog/inzicht-je-missing-data/)

- [Omgaan met missing data](https://www.dpa.nl/it/blog/methoden-om-dataset-compleet-maken/)

---

### Working papers

- Schouten, R.M. (2017) [The effect of missing data mechanisms on the validity of regression coefficients]()

- Schouten, R.M. (2017) Evaluation of missing data methods in the crossvalidation framework: MSE and R^2

---

### Software development

- `ampute` in R-package **mice**

```{r, eval = FALSE}
require(mice)
?ampute
```

Ampute does this and this. An extensive vignette that explains the method and how to use the function can be found [hier is de link](https://github.com/RianneSchouten/Amputation_with_Ampute/tree/master/Vignette).

- `parlMICE` 

```{r, eval = FALSE}
#something
```

`parlMICE` is a wrapper function, which I developed together with Gerko Vink. It can be used when multiple imputations are done with `mice` from R-package **mice** but with large m. Explanation of the method of multiple imputation can be found [here](https://www.jstatsoft.org/article/view/v045i03). A vignette explaining the value and use of `parlMICE` is available [here](https://github.com/RianneSchouten/parlMICE).

- `MCAR_Tests` in Python

I developed two tests to check whether an incomplete dataset has a MCAR mechanisms. For each pair of variables, t-tests can be performed with `t-test function`. Little's MCAR test is implemented in `mcar test`. The code can be found [here]().

---

### Talks and conferences

- useR!2017 Brussels

My [presentation](C:/Users/SCHOUT07/Documents/Presentaties/UseR!2017/Multivariate_Amputation_Rianne_Schouten.pdf) discussed the generation of missing values with `ampute`.

- Amst-R-Dam 2017

A bunch of R-fanatics listened to me talking about amputation and how to use `ampute` to do this. The presentation can be found [here](C:/Users/SCHOUT07/Documents/Presentaties/Amst-R-Dam/Multivariate_Amputation_Rianne_Schouten.pdf).

- eScience Symposium [planned]

- Women in Technology [planned]

---

### Publications

- Schouten, R.M., Lugtig, P.J., Brand, J.P.L. and Vink, G. (2017). Generating missing values for simulation
purposes: A multivariate amputation procedure. [Under review at Journal of Statistical Computation and Simulation]

- Schouten, R.M. and Vink, G. (2017). The dance of the mechanisms: how observed information influences the validity of missingness assumptions. [Under review at Sociological Methods \& Research]

- de Vries, C. P., Schouten, R. M., van der Kuur, J., Gottardi, L., & Akamatsu, H. (2016, July). [Microcalorimeter pulse analysis by means of principle component decomposition](link). In SPIE Astronomical Telescopes+ Instrumentation (pp. 99055I-99055I). International Society for Optics and Photonics.

- Kappen, I.F.P.M., Bittermann, G.K.P., Schouten, R.M., Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. and Breugem, C.C. (2017). [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis](C:/Users/SCHOUT07/Documents/Papers/Cephalometric_Analysis/CephalometricAnalysis.pdf). Clinical Oral Investigations, 21, pp. 1801-1810

---

### Contact

- [Curriculum Vitae](C:/Users/SCHOUT07/Documents/Papieren/CV/DPAIT_RianneSchouten_DataScientist_ned.docx)

- Email: r.m.schouten\@uu.nl; rianne.schouten\@dpa.nl

- Telefoon: +31(0)640640516

- Zie ook mijn [Github](https://github.com/RianneSchouten) of [LinkedIn](https://www.linkedin.com/in/rianne-schouten-5bb29491/)

---
