---
title: "Rianne Schouten"
author: |
  | PhD Candidate / Missing Data Specialist / Junior Data Scientist
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

---

### Introduction and affiliations

When you get to know me, you will recognize an enthusiastic and motivated person who is always eager to learn. I have an open and direct personality, and some self-deprecating humor. Working with me never becomes boring, since I continuously make new plans or I will make some weird (but funny) statements. First and foremost, I intent to deliver work with a more than good quality. 

As a PhD student, I work under supervision of [Gerko Vink](https://www.gerkovink.com) and [Stef van Buuren](https://github.com/stefvanbuuren) at the Department of Methodology and Statistics, University of Utrecht, the Netherlands. The past year, I developed and implemented multivariate amputation procedure `ampute` (i.e. a way to generate sophisticated missingness in complete datasets). With the availability of `ampute` in R-package `mice`, it has become possible to perform extensive simulation studies to evaluate the effect of missing data methods. My PhD research proposal [About the evaluation of missing data methodologies](pdfs/ResearchProposal_RianneSchouten.pdf) gives some specifics about my current work. 

Furthermore, I work as a Missing Data Specialist at DPA Professionals, Business Unit IT. I advise companies about missing data problems and possible solutions. With my work, I contribute to the creation of correct and rightful (statistical) models. At DPA, I am also part of the [Data Science Excellence Program](http://www.dpadatascience.nl). This means that I receive training from senior data scientists at [Big Data Republic](https://www.bigdatarepublic.nl/) and learn how to apply machine learning techniques to answer practical questions. 

---

### Publications

- Schouten, R.M., Lugtig, P.J. and Vink, G. (2017). [Generating missing values for simulation purposes: A multivariate amputation procedure](https://rianneschouten.github.io/Amputation_with_Ampute/Manuscript%20article/Manuscript.pdf). [Under review at Journal of Statistical Computation and Simulation]

- Schouten, R.M. and Vink, G. (2017). The dance of the mechanisms: how observed information influences the validity of missingness assumptions. [Under review at Sociological Methods \& Research]

- Kappen, I.F.P.M., Bittermann, G.K.P., Schouten, R.M., Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. and Breugem, C.C. (2017). [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis](pdfs/CephalometricAnalysis.pdf). Clinical Oral Investigations, 21, pp. 1801-1810.

- de Vries, C.P., Schouten, R.M., Van der Kuur, J., Gottardi, L., and Akamatsu, H. (2016, July). [Microcalorimeter pulse analysis by means of principle component decomposition](pdfs/Microcalorimeter_Pulse_Analysis.pdf). In SPIE Astronomical Telescopes+ Instrumentation (pp. 99055I-99055I). International Society for Optics and Photonics.

---

### Working papers

- Schouten, R.M. (2017) [The effect of missing data mechanisms on the validity of regression coefficients](pdfs/Regression_coefficients_Rianne_Schouten.pdf).

- Schouten, R.M. (2017) Evaluation of missing data methods in the data science framework: [prediction accuracy](pdfs/Prediction_accuracy.pdf) and [imputation accuracy](pdfs/Imputation_accuracy.pdf). 

---

### Software development

#### 1. `ampute` in R-package **mice**

```{r, eval = FALSE}
require(mice)
?ampute
```

R-function `ampute` is the implementation of a [multivariate amputation procedure](https://rianneschouten.github.io/Amputation_with_Ampute/Manuscript%20article/Manuscript.pdf). With `ampute`, it is straightforward to generate missing values in multiple variables, with different missing data proportions and varying underlying missingness mechanisms. An extensive [vignette](https://rianneschouten.github.io/Amputation_with_Ampute/Vignette/ampute.html) is available.

#### 2. `parlMICE` 

```{r, eval = FALSE}
require(devtools)
source_url('https://raw.githubusercontent.com/RianneSchouten/parlMICE/master/parlMICE.R')
```

With large datasets or with higher number of imputations, performing multiple imputation with `mice` from R-package **mice** might take a long time. As a solution, I created wrapper function `parlMICE` in cooperation with Gerko Vink. With `parlMICE`, the imputation procedure can be run in parallel. A [vignette](https://github.com/RianneSchouten/parlMICE) explaining the value and use of `parlMICE` is available.

#### 3. `pymice`: mice in Python

Although many missing data functions exist in `R`, the collection of missing data functionS in `Python` is quite poor. I created the start of a new package which will eventually consist of three parts: missing data exploration, multivariate amputation and multiple imputation. 

At the moment, the package contains class `McarTests`, which consists of two functions to inspect whether the nonresponse has a MCAR missingness mechanism. Little's MCAR test is implemented in `mcar_test` and for each pair of variables, t-tests can be performed with function `mcar_t_tests`. The code can be found in my Github repository called [`pymice`](https://github.com/RianneSchouten/pymice).

---

### Blogposts

- [Inzicht in je missing data](https://www.dpadatascience.nl/blog/inzicht-missing-data/)

- [Omgaan met missing data](https://www.dpadatascience.nl/blog/omgaan-missing-data/)

---

### Invited talks

- Data Science Hackathon

A monthly organized [data science hackathon](https://www.meetup.com/nl-NL/Eindhoven-Data-Science-Meetup/events/245332474/?eventId=245332474&chapter_analytics_code=UA-80541826-1) will be about missing data problems. I participate in the organization of the hackathon, will provide the attendants with the necessary knowledge and will give them feedback on their approach. 

- useR!2017 Brussels

My presentation discussed the generation of missing values with `ampute`. A [summary](https://user2017.sched.com/event/AxsS?iframe=no) can be found on the website of the conference. 

- Amst-R-Dam 2017

A bunch of R-enthusiasts listened to me talking about amputation and how to use `ampute` to do this. 

---

### Attended conferences

- Women in Technology [8 and 9 November, 2017]

The [Women in Technology](http://join.europeanwomenintech.com/technology/?gclid=CjwKEAjwruPNBRCKkbL9zqKcrHwSJABGDVyIbRE8y7tfRU6cobtIyGOBlg-HI1dfOAbqvssG35vhnxoC0mHw_wcB) Series enables the sector to connect, learn and act on gender diversity by sharing the experiences of industry leaders and developing women's skills, both soft and technical. I wrote a [blogpost](https://www.linkedin.com/pulse/i-want-go-where-am-invited-included-rianne-schouten/) to share my experiences. 

- useR!2017 Brussels [4 - 7 July, 2017]

The annual [useR!](https://user2017.brussels/) conference is the main meeting of the international R user and developer community.

---

### Contact

- Curriculum Vitae in [Dutch](pdfs/Rianne_Schouten_CV_DPA_IT_Nederlands.pdf) or [English](pdfs/Rianne_Schouten_CV_DPA_IT_English.pdf).

- Email: r.m.schouten\@uu.nl; rianne.schouten\@dpa.nl

- Telephone: +31(0)640640516

- Also see my [Github](https://github.com/RianneSchouten) or [LinkedIn](https://www.linkedin.com/in/rianne-schouten-5bb29491/)

---
