---
title: "Rianne Schouten"
subtitle: "PhD Candidate"
author: |
  | Pattern Mining | Exceptional Model Mining | Missing Data 
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

---

### Introduction and affiliations

My name is Rianne and I develop data analysis methods.

I am a PhD Candidate at Eindhoven University of Technology and study the Exceptional Model Mining framework. We work on developing solutions for heterogeneous datasets with time-dependent attributes. My supervisors are [dr. Wouter Duivesteijn](http://wwwis.win.tue.nl/~wouter/) and [prof. Mykola Pechenizkiy](https://www.win.tue.nl/~mpechen/). 

Additionally, I work on Missing Data methods, particularly the process of generating missing values in complete datasets. Under supervision of [prof. Stef van Buuren](https://scholar.google.nl/citations?user=_3y5C0UAAAAJ&hl=nl) and [dr. Gerko Vink](https://www.gerkovink.com/), I developed a multivariate amputation procedure and implemented the method in an R-function ([`ampute`](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html) in package [`mice`](https://github.com/stefvanbuuren/mice)) and python module ([`pyampute`](https://rianneschouten.github.io/pyampute/build/html/index.html)).

I find it important that my work can be used for real-world problems (e.g. in health care) and enjoy collaborating on projects. Please contact me if you have questions or seek opportunities to work together. 

Below you will find overviews of and links to [Publications], [Awards], [Software development], [Teaching activities], [Supervision], [Review activities], [Presentations], [Personal development], and [Contact details]

---

### Publications

- [In press] Verhaegh, Ruben F.A., Kiezebrink, Jacco J.E., Nusteling, Frank, Rio, Arnaud W.A., Bendicsek, Marton B., Duivesteijn, Wouter & Schouten, Rianne M. (2022) [A Clustering-inspired Quality Measure for Exceptional Preferences Mining — Design Choices and Consequences](pdfs/2022_Verhaegh_EPM.pdf). In: Proceedings of the International Conference on Discovery Science (DS).

- Schouten, Rianne. M., Zamanzadeh, Davina & Singh, Prabhant (2022) [Pyampute: a Python library for data amputation](https://rianneschouten.github.io/pyampute/build/html/index.html). Zenodo.  [https://doi.org/10.25080/majora-212e5952-03e](https://doi.org/10.25080/majora-212e5952-03e).

- Van der Haar, J.F., Nagelkerken, S.C., Smit, I.G., van Straaten, K., Tack, J.A., Schouten, R.M. & Duivesteijn, W. (2022) [Efficient Subgroup Discovery Through Auto-Encoding.](pdfs/2022_vdHaar_AE.pdf) In: Proceedings of the 20th International Symposium on Intelligent Data Analysis (IDA), pp. 327-340.

- Schouten, R.M., Duivesteijn, W. & Pechenizkiy, M. (2022) [Exceptional Model Mining for Repeated Cross-Sectional Data (EMM-RCS)](pdfs/2022_Schouten_EMMRCS.pdf). In: Proceedings of SIAM International Conference on Data Mining (SDM), pp. 585-593.

- Schouten, R.M., Bueno, M.L.P., Duivesteijn, W. & Pechenizkiy, M. (2022) [Mining Sequences with Exceptional Transition Behaviour of Varying Order using Quality Measures based on Information-Theoretic Scoring Functions.](pdfs/MiningExceptionalSequences_DAMI_2022.pdf) Data Mining and Knowledge Discovery (DAMI), 36: 379-413.

- IJsselhof R, Duchateau S, Schouten R, Slieker M, Hazekamp M & Schoof P. [Long-Term Follow-Up of Pericardium
for the Ventricular Component in Atrioventricular Septal Defect Repair.](https://pubmed.ncbi.nlm.nih.gov/33164688/) World Journal for Pediatric and Congenital Heart Surgery, 11(6): 742-747.

- IJsselhof R.J., Duchateau S.D.R., Schouten R.M., Freund, M.W., Heuser, J., Fejzic, Z., Haas, F., Schoof, P.H. & Slieker, M.G. (2019) [Follow-up After Biventricular Repair of the Hypoplastic Left Heart Complex.](https://academic.oup.com/ejcts/article-abstract/57/4/644/5607276) European Journal of Cardiothoracic Surgery, 57(4): 644-651.

- Rianne R.M., Lugtig, P. & Vink, G. (2018) [Generating missing values for simulation purposes: A multivariate amputation procedure.](https://www.tandfonline.com/doi/full/10.1080/00949655.2018.1491577) Journal of Statistical Computation and Simulation, 88(15): 1909-1930.

- Schouten, R.M. and Vink, G. (2021) [The dance of the mechanisms: How observed information influences the validity of missingness assumptions.](https://journals.sagepub.com/doi/full/10.1177/0049124118799376) Sociological Methods \& Research, 50(3): 1243-1258.

- Kappen, I.F.P.M., Bittermann, G.K.P., Schouten, R.M., Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. & Breugem, C.C. (2017) [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis.](https://www.researchgate.net/publication/308181181_Long-term_mid-facial_growth_of_patients_with_a_unilateral_complete_cleft_of_lip_alveolus_and_palate_treated_by_two-stage_palatoplasty_cephalometric_analysis) Clinical Oral Investigations, 21: 1801-1810.

- de Vries, C.P., Schouten, R.M., Van der Kuur, J., Gottardi, L., & Akamatsu, H. (2016) [Microcalorimeter pulse analysis by means of principle component decomposition.](https://arxiv.org/abs/1607.00834) Proceedings SPIE 9905, Space TElescopes and Instrumentation 2016: Ultraviolet to Gamma Ray, 99055v. DOI: 10.1117/12.2231627

---

### Awards

1. Performance bonus for 2021

> "Beside your overall excellent performance in your PhD research and EDIC project execution, you did an excellent job in project management of EDIC, and in setting up new successful collaborations. You helped a lot with the Research Topics in Data Mining course, and supervision of students."

2. Pluim for Excellent course evaluation *Research Topics in Data Mining* 2021\/2022

---

### Software development

#### 1. `ampute` in R-package **mice**

```{r, eval = FALSE}
library(mice)
?ampute
```

R-function `ampute` is the implementation of a multivariate amputation procedure: a method for generating missing data in complete datasets. With `ampute`, it is straightforward to generate missing values in multiple variables, with different missing data proportions and varying underlying missingness mechanisms. Read [the article](pdfs/ArticleAmpute.pdf) or [the vignette](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html) to learn more.

#### 2. `parlMICE` 

For large datasets or when you want to impute with a large number of imputations, multiple imputation with `mice` in R-package **mice** may have a long run time. As a solution, Gerko Vink and I created wrapper function `parlMICE`, which allows for a parallel run of `mice`.

The function is now part of package **mice** under the name `parlmice`.

```{r, eval = FALSE}
library(mice)
?parlmice
```

All information can be found [in the github repo](https://github.com/RianneSchouten/parlMICE) or in the [vignette](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html).

#### 3. `pyampute`: the first Python library for data amputation

Library `pyampute` provides the multivariate amputation methodology to the Python community, and it does more. It has improved default settings, allows for a combined MAR+MNAR mechanism, for custom probability functions and since it is compatible with scikit-learn's `fit` and `transform` paradigm, seamless integration in data processing pipelines becomes easy. 

Find plenty of examples in [`pyampute`'s documentation](https://rianneschouten.github.io/pyampute/build/html/index.html). Davina's presentation at SciPy22 can be found [here](https://www.youtube.com/watch?v=iI0gx52ucMA&t=25560s).

Install using pip or from source:

```
pip install pyampute
```

```
git clone https://github.com/RianneSchouten/pyampute.git
pip install ./pyampute
```

--- 

### Teaching activities

During my time at TU\/e, I have taught in the following courses.

|Year|Course|Level|Activities|
|----|------|------|------------|
|20\/21|2AMM20 Foundations of Data Mining|MSc|Taught 2 lectures about Missing Data, provided answers in weekly Q\&A, developed practice questions for exam, administrative activities|
|21\/22|2AMM20 Research Topics in Data Mining|MSc|Taught 2 lectures about Missing Data, supervised 7 groups of students during research project|
|22\/23|2AMM20 Research Topics in Data Mining|MSc||

During my time at UU, I have taught lectures in 2 summerschool courses in 2015 and 2016. The courses are at an Advanced Master level: [Survey Research: Design, Implementation and Data Processing](https://utrechtsummerschool.nl/courses/social-sciences/survey-research-design-implementation-and-data-processing) and [Survey Research: Statistical Analysis and Estimation](https://utrechtsummerschool.nl/courses/social-sciences/survey-research-statistical-analysis-and-estimation).

---

### Supervision activities

During my time at TU\/e, I have supervised the following students. 

|Year|Student|Type|Topic|
|----|------|------|------------|
|20\/21|Bart van Dooren|MSc Thesis with Philips|Predicting Cardiovascular Risk with Objective Physical Activity Measurements|
|21\/22|Mats Verbraak|Research Proposal|Handling Missing Data in the Prediction Domain using Multiple Imputation|
|21\/22|Isabel van den Heuvel|Research Proposal|Equivalence Testing for Developing Fair Machine Learning Algorithms|
|21\/22|Varun Kamat|Internship at Signify|Recommendation Tool for Component Database|
|21\/22|Mika van Loon|MSc Thesis||
---

### Review activities

With help from my supervisors, I have reviewed for Data Mining and Knowledge Discovery (DAMI) and the journal track of ECMLPKDD 2020 (Machine Learning).

---

### Presentations

|Year|Occasion|Type|Topic|Link to materials|
|----|------|------|------------|----|
|2021|Neglected Assumptions in Causal Inference (NACI) workshop at ICML|Contribution|Understanding the Role of Prognostic Factors and Effect Modifiers in Heterogeneity of Treatment Effect using a Within-Subjects Analysis of Variance|[link](pdfs/NACI_Poster_RMSchouten.pdf)|
|2021|EAISI Eindhoven|Contribution to seminar|Towards a better understanding of exceptional lifestyle behaviour||
|2021|ECMLPKDD|Poster at conference|Mining Sequences with Exceptional Transition Behaviour of Varying Order|[link](pdfs/PosterSatelliteEvent_RMSchoutenetal.pdf)|
|2019|Workshop R-Ladies Amsterdam|Invited presentation at seminar|Developed and presented a workshop about analysis of missing values, evaluation and implementation of missing data methods|[link](https://github.com/RianneSchouten/workshops/tree/master/R-Ladies%20Workshop)|
|2018|ICT Open|Contribution to 1-day conference|Handling Missing Data in Data Science|[link](pdfs/ictopen2018_RianneSchouten.pdf)|
|2018|European Women in Technology|Masterclass at conference|Dealing with missing data in `R`: Amputation or Imputation?|[presentation and exercises](https://github.com/RianneSchouten/workshops/tree/master/WomenInTechnology)|
|2018|sat-R-Day|Contribution to 1-day conference|Missing data|[link](pdfs/SatRdayRianneSchouten.pdf)|
|2018|Data Science Hackathon|By invitation|Developed and lead a missing data challenge|[link](pdfs/Data_Science_Hackathon_Rianne_Schouten.pdf)|
|2017|Amst-R-Dam|Contribution to seminar|How to use R-function `ampute` to generate missing values in complete datasets|[article](pdfs/ArticleAmpute.pdf) and [documentation](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html)|
|2017|UseR!2017|Contribution to conference|Introduction to multivariate amputation with `ampute`||

---

### Personal development

To develop myself as a teacher, I am participating in the University Teaching Qualification (UTQ) Training Program.

Other courses that I participated in are Project management (PROOF TU/e 2020), Introduction to Process Mining (Fluxicon, 2020), Learning and Reasoning (SIKS, 2021) and Communication Styles (PROOF TU/e 2021).


---

### Contact details

- Email: r.m.schouten\@tue.nl

- Github: [/RianneSchouten](https://github.com/RianneSchouten)

- [LinkedIn](https://www.linkedin.com/in/rianne-schouten-5bb29491/)

- Twitter: [missD_ta](https://twitter.com/missD_ta)

- [Google Scholar](https://scholar.google.nl/citations?view_op=list_works&hl=nl&user=xC56bhMAAAAJ&gmla=AJsN-F4IWn_9Vw9hH7hPN1zIMfNadBL6mvmu-rPpRS6p-4Enc7fR6GrOMYyN2QEr21D5Y3Uhjtj0HPT3I7qsh3MVE1v2L8MRfjsk7W1Qop4mDunpnTRDeMYF0DmoRYmja6b7GvnmC4R3)

- [Research Gate](https://www.researchgate.net/profile/Rianne_Schouten/publications)

- [Curriculum Vitae](pdfs/Rianne_Schouten_Resume.pdf)

