---
title: "Rianne Schouten"
author: |
  | PhD Candidate | Exceptional Model Mining | Missing Data | Causal Inference
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

---

### Introduction and affiliations

My name is Rianne and I analyze analysis methods. 

From 2016 - 2019 I studied the effect of missing data on analysis outcomes with [prof dr. Stef van Buuren](https://scholar.google.nl/citations?user=_3y5C0UAAAAJ&hl=nl) and [dr. Gerko Vink](https://www.gerkovink.com/). During this time, I co-developed a procedure to generate missing data in complete datasets and implemented the method in R-function [`ampute`](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html). It is part of R-package [`mice`](https://github.com/stefvanbuuren/mice).

Currently, I work in the Data Mining group at Eindhoven University of Technology and study the Exceptional Model Mining framework. We work on developing solutions for heterogeneous datasets with time-dependent attributes. My supervisors are [dr. Wouter Duivesteijn](http://wwwis.win.tue.nl/~wouter/) and [prof. dr. Mykola Pechenizkiy](https://www.win.tue.nl/~mpechen/). 

I find it important that my work can be used for real world problems. Especially the health care domain has my interest and I joined as a statistical advisor in several medical studies. From both a missing data and medical point of view, I am interested in Causal Inference as well.

If you have questions regarding exceptional model mining, missing data, causal inference, statistical techniques, simulation studies or anything else related to data analysis, feel free to contact me (see below).

---

### Publications

- IJsselhof R, Duchateau S, Schouten R, Slieker M, Hazekamp M & Schoof P. [Long-Term Follow-Up of Pericardium
for the Ventricular Component in Atrioventricular Septal Defect Repair.](https://pubmed.ncbi.nlm.nih.gov/33164688/) World Journal for Pediatric and Congenital Heart Surgery, 11(6): 742-747.

- IJsselhof R.J., Duchateau S.D.R., Schouten R.M., Freund, M.W., Heuser, J., Fejzic, Z., Haas, F., Schoof, P.H. & Slieker, M.G. (2019) [Follow-up After Biventricular Repair of the Hypoplastic Left Heart Complex.](https://academic.oup.com/ejcts/article-abstract/57/4/644/5607276) European Journal of Cardiothoracic Surgery, 57(4): 644-651.

- Rianne R.M., Lugtig, P. & Vink, G. (2018) [Generating missing values for simulation purposes: A multivariate amputation procedure.](https://www.tandfonline.com/doi/full/10.1080/00949655.2018.1491577) Journal of Statistical Computation and Simulation, 88(15): 1909-1930. DOI: 10.1080/00949655.2018.1491577

- Schouten, R.M. and Vink, G. (2018) [The dance of the mechanisms: How observed information influences the validity of missingness assumptions.](https://journals.sagepub.com/doi/full/10.1177/0049124118799376) Sociological Methods \& Research. DOI: 10.1177/0049124118799376

- Kappen, I.F.P.M., Bittermann, G.K.P., Schouten, R.M., Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. & Breugem, C.C. (2017) [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis.](https://www.researchgate.net/publication/308181181_Long-term_mid-facial_growth_of_patients_with_a_unilateral_complete_cleft_of_lip_alveolus_and_palate_treated_by_two-stage_palatoplasty_cephalometric_analysis) Clinical Oral Investigations, 21: 1801-1810.

- de Vries, C.P., Schouten, R.M., Van der Kuur, J., Gottardi, L., & Akamatsu, H. (2016) [Microcalorimeter pulse analysis by means of principle component decomposition.](https://arxiv.org/abs/1607.00834) Proceedings SPIE 9905, Space TElescopes and Instrumentation 2016: Ultraviolet to Gamma Ray, 99055v. DOI: 10.1117/12.2231627

---

### Software development

#### 1. `ampute` in R-package **mice**

```{r, eval = FALSE}
library(mice)
?ampute
```

R-function `ampute` is the implementation of a multivariate amputation procedure: a method to generate missing data in complete datasets. With `ampute`, it is straightforward to generate missing values in multiple variables, with different missing data proportions and varying underlying missingness mechanisms. Read [the article](pdfs/ArticleAmpute.pdf) or [the vignette](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html) to learn more.

#### 2. `parlMICE` 

For large datasets or with a high number of imputations, multiple imputation with `mice` in R-package **mice** may have a long running time. As a solution, Gerko Vink and I created wrapper function `parlMICE`, which allows for a parallel run of `mice`.

The function is now part of package **mice** under the name `parlmice`.

```{r, eval = FALSE}
library(mice)
?parlmice
```

All information can be found [in the github repo](https://github.com/RianneSchouten/parlMICE) or in the [vignette](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html).

---

### Invited presentations

##### * Workshop R-Ladies Amsterdam, October 2019

[Find the materials here](https://github.com/RianneSchouten/workshops/tree/master/R-Ladies%20Workshop)

The workshop covers: 

1.	Missing Values Analysis
2.	Evaluating Missing Data methods
3.	Implementing Missing Data Methods

##### * Masterclass @ European Women in Technology, November 2018

Dealing with missing data in R: Amputation or Imputation? Find here the [presentation and exercises](https://github.com/RianneSchouten/workshops/tree/master/WomenInTechnology).

##### * sat-R-Day, September 2018

An explanation of missing data mechanisms, missing data patterns and how to 
generate them with `ampute`. Check the [slides](pdfs/SatRdayRianneSchouten.pdf)

##### * ICT Open, March 2018

Handling Missing Data in Data Science: Simulating the effects of missing data methods and how to present the results in an interactive plot with Github Pages. Read the [abstract](pdfs/abstract_rianne_schouten.pdf) or check the [slides](pdfs/ictopen2018_RianneSchouten.pdf)

##### * Data Science Hackathon, January 2018

My presentation can be found [here](pdfs/Data_Science_Hackathon_Rianne_Schouten.pdf).

##### * useR!2017 Brussels, July 2017

My presentation discussed the generation of missing values with `ampute`. A [summary](https://user2017.sched.com/event/AxsS?iframe=no) can be found on the website of the conference. 

##### * Amst-R-Dam 2017

A presentation about amputation and how to use `ampute` to do this. 

---

### Contact

- Email: r.m.schouten\@tue.nl

- Github: [/RianneSchouten](https://github.com/RianneSchouten)

- [LinkedIn](https://www.linkedin.com/in/rianne-schouten-5bb29491/)

- Twitter: [missD_ta](https://twitter.com/missD_ta)

- [Google Scholar](https://scholar.google.nl/citations?view_op=list_works&hl=nl&user=xC56bhMAAAAJ&gmla=AJsN-F4IWn_9Vw9hH7hPN1zIMfNadBL6mvmu-rPpRS6p-4Enc7fR6GrOMYyN2QEr21D5Y3Uhjtj0HPT3I7qsh3MVE1v2L8MRfjsk7W1Qop4mDunpnTRDeMYF0DmoRYmja6b7GvnmC4R3)

- [Research Gate](https://www.researchgate.net/profile/Rianne_Schouten/publications)

- [Curriculum Vitae](pdfs/Rianne_Schouten_Resume.pdf)

