---
title: "Rianne Schouten"
author: |
  | PhD Candidate | Exceptional Model Mining | Missing Data
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

---

### Introduction and affiliations

My name is Rianne and I analyze analysis methods. From 2016 - 2019 I studied the effect of missing data on analysis outcomes. During this time, I co-developed a procedure to generate missing data in complete datasets and implemented the method in R-function `ampute`. It is part of R-package `mice`.

Currently, I work in the Data Mining group at the TUe and study the Exceptional Model Mining framework. We work on developing solutions for heterogeneous datasets with time-dependent attributes. My supervisors are dr. Wouter Duivesteijn and prof. dr. Mykola Pechenizkiy. 

I find it important that my work can be used for real world problems. Especially the health care domain has my interest and I co-operated in several medical studies as a statistical advisor. 

If you have questions regarding exceptional model mining, missing data, statistical techniques, simulation studies or anything else related to data analysis, feel free to contact me.

---

### Publications

- **Research accepted**: IJsselhof R, Duchateau S, Schouten R, Slieker M, Hazekamp M & Schoof P. Untreated autologous pericardium for ventricular septal defect closure in atrioventricular septal defect repair. World Journal for Pediatric and Congenital Heart Surgery. Date of submission: 15-01-2020.

- **In press**: IJsselhof R, Duchateau S, Schouten R, et al. (2019) Follow-up After Biventricular Repair of the Hypoplastic Left Heart Complex. European Journal of Cardiothoracic Surgery.

- Rianne Margaretha Schouten, Peter Lugtig & Gerko Vink (2018): [Generating missing values for simulation purposes: A multivariate amputation procedure](pdfs/ArticleAmpute.pdf). Journal of Statistical Computation and Simulation, 88(15): 1909-1930. DOI: 10.1080/00949655.2018.1491577

- Schouten, R.M. and Vink, G. (2018). [The dance of the mechanisms: How observed information influences the validity of missingness assumptions](pdfs/TheDance.pdf). Sociological Methods \& Research. DOI: 10.1177/0049124118799376

- Kappen, I.F.P.M., Bittermann, G.K.P., Schouten, R.M., Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. and Breugem, C.C. (2017). [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis](pdfs/CephalometricAnalysis.pdf). Clinical Oral Investigations, 21, pp. 1801-1810.

- de Vries, C.P., Schouten, R.M., Van der Kuur, J., Gottardi, L., and Akamatsu, H. (2016, July). [Microcalorimeter pulse analysis by means of principle component decomposition](pdfs/Microcalorimeter_Pulse_Analysis.pdf). In SPIE Astronomical Telescopes+ Instrumentation (pp. 99055I-99055I). International Society for Optics and Photonics.

---

### Software development

#### 1. `ampute` in R-package **mice**

```{r, eval = FALSE}
require(mice)
?ampute
```

R-function `ampute` is the implementation of a multivariate amputation procedure: a method to generate missing data in complete datasets. With `ampute`, it is straightforward to generate missing values in multiple variables, with different missing data proportions and varying underlying missingness mechanisms. Read [the article](pdfs/ArticleAmpute.pdf) or [the vignette](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html) to learn more.

#### 2. `parlMICE` 

```{r, eval = FALSE}
require(devtools)
source_url('https://raw.githubusercontent.com/RianneSchouten/parlMICE/master/parlMICE.R')
```

With large datasets or with a large number of imputations, multiple imputation with `mice` in R-package **mice** may have a long running time. As a solution, Gerko Vink and I created wrapper function `parlMICE`. With `parlMICE`, the imputation procedure can be run in parallel. Find all information [in the github repo](https://github.com/RianneSchouten/parlMICE) or read the [vignette](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html).

---

### Workshops and presentations

##### * Workshop R-Ladies Amsterdam, October 2019

[Find the materials here](https://github.com/RianneSchouten/workshops/tree/master/R-Ladies%20Workshop)

The workshop covers: 

1.	Missing Values Analysis
2.	Evaluating Missing Data methods
3.	Implementing Missing Data Methods

##### * Masterclass @ European Women in Technology, November 2018

Dealing with missing data in R: Amputation or Imputation? Find here the [presentation and exercises](https://github.com/RianneSchouten/workshops/tree/master/WomenInTechnology).

##### * satRDay, September 2018

An explanation of missing data mechanisms, missing data patterns and how to 
generate them with `ampute`. Check the [slides](pdfs/SatRdayRianneSchouten.pdf)

##### * ICT Open, March 2018

Handling Missing Data in Data Science: Simulating the effects of missing data methods and how to present the results in an interactive plot with Github Pages. Read the [abstract](pdfs/abstract_rianne_schouten.pdf) or check the [slides](pdfs/ictopen2018_RianneSchouten.pdf)

##### * Data Science Hackathon, January 2018

My presentation can be found [here](pdfs/Data_Science_Hackathon_Rianne_Schouten.pdf).

##### * useR!2017 Brussels, July 2017

My presentation discussed the generation of missing values with `ampute`. A [summary](https://user2017.sched.com/event/AxsS?iframe=no) can be found on the website of the conference. 

##### * Amst-R-Dam

A presentation about amputation and how to use `ampute` to do this. 

---

### Other attended conferences 

##### * ICML 2020

##### * Big Data Expo, September 2018

##### * Women in Technology, November 2017

The [Women in Technology](http://join.europeanwomenintech.com/technology/?gclid=CjwKEAjwruPNBRCKkbL9zqKcrHwSJABGDVyIbRE8y7tfRU6cobtIyGOBlg-HI1dfOAbqvssG35vhnxoC0mHw_wcB) Series enables the sector to connect, learn and act on gender diversity by sharing the experiences of industry leaders and developing women's skills, both soft and technical. I wrote a [blogpost](https://www.linkedin.com/pulse/i-want-go-where-am-invited-included-rianne-schouten/) to share my experiences. 

##### * CWI in Bedrijf, May 2017

---

### Contact

- Email: r.m.schouten\@tue.nl

- Also see my [Github](https://github.com/RianneSchouten) or [LinkedIn](https://www.linkedin.com/in/rianne-schouten-5bb29491/)

- Curriculum Vitae in [English](pdfs/Rianne_Schouten_Resume.pdf).

