---
title: "Rianne M. Schouten"
subtitle: "Postdoctoral researcher @ TU/e"
author: ""
date: ""
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    theme: spacelab
    highlight: tango
---

<script>
   $(document).ready(function() {
     $head = $('#header');
     $head.prepend('<img src=\"photo.jpg\" style=\"float: right;width: 250px;\"/>')
   });
</script>

---

[Google Scholar](https://scholar.google.nl/citations?view_op=list_works&hl=nl&user=xC56bhMAAAAJ&gmla=AJsN-F4IWn_9Vw9hH7hPN1zIMfNadBL6mvmu-rPpRS6p-4Enc7fR6GrOMYyN2QEr21D5Y3Uhjtj0HPT3I7qsh3MVE1v2L8MRfjsk7W1Qop4mDunpnTRDeMYF0DmoRYmja6b7GvnmC4R3)

[LinkedIn](https://www.linkedin.com/in/rianne-schouten-phd-5bb29491/)

[Github](https://github.com/RianneSchouten)

E-mail: [r.m.schouten@tue.nl](mailto:r.m.schouten@tue.nl)

Please contact me for my research and teaching statements. 

---

### Introduction

**Methods**: Exceptional Model Mining | Local Pattern Mining | Subgroup Discovery 

**Type of Data**: Non-IID | Hierarchical | Sequential | Time Series | Longitudinal 

**Application domains**: Personalized Learning Analytics | e-Health | Medicine 

My research revolves around developing local pattern mining methods for extracting societal relevant, interpretable patterns from non-IID data, such as discovering subgroups of patients with exceptional responses to treatment.My next steps include developing extensions for multi-modal data, for online learning and for integrating domain knowledge and user feedback.

---

### Affiliations

[See here for complete Curriculum Vitae.](pdfs/ResumeRianneSchouten.pdf)

**2025 - present**: **Post-doc** at TU/e. 

**2020 - 2024**: **Ph.D. candidate** at TU/e.

Supervisors: [Prof. Mykola Pechenizkiy](https://www.win.tue.nl/~mpechen/), [Dr. Wouter Duivesteijn](http://wwwis.win.tue.nl/~wouter/).

Topic: Exceptional Model Mining for Hierarchical Data. 

**2017 - 2019**: **Missing Data Researcher** at Utrecht University.

Supervisors: [Prof. dr. Stef van Buuren](https://scholar.google.nl/citations?user=_3y5C0UAAAAJ&hl=nl), [Dr. Gerko Vink](https://www.gerkovink.com/).

**2016**: **Staff Associate** of Professor Andrew Gelman at Columbia University in the City of New York, US. Duration: 2 months.

**2015**: **Intern** at SRON, Dutch Institute of Space Research, the Netherlands. Duration: 4 months. 

---

### Publications

[See Google Scholar.](https://scholar.google.nl/citations?view_op=list_works&hl=nl&user=xC56bhMAAAAJ&gmla=AJsN-F4IWn_9Vw9hH7hPN1zIMfNadBL6mvmu-rPpRS6p-4Enc7fR6GrOMYyN2QEr21D5Y3Uhjtj0HPT3I7qsh3MVE1v2L8MRfjsk7W1Qop4mDunpnTRDeMYF0DmoRYmja6b7GvnmC4R3)

- van den Biggelaar, L., **Schouten, R.M.**, de Bie, A., Bouwman, A, \& Duivesteijn. W. Characterizing the Risk of Atrial Fibrillation in Cardiac Patients with Exceptional Electrocardiogram Phenotypes. Accepted for publication at KDD25.

- **Schouten, R.M.** (2025) [Exceptional Model Mining for Hierarchical Data](pdfs/EMMHD.pdf). PhD thesis.

- **Schouten, R.M.** (2024) [On the role of prognostic factors and effect modifiers in structural causal models](pdfs/NeurIPS_CRL_RianneSchouten_nr42.pdf). Accepted for presentation at Causal Representation Learning Workshop NeurIPS. [See here for the paper!](pdfs/NeurIPSCRL_RianneSchouten_nr42.pdf)

- Van den Berg, N. T., Broekgaarden, B. O., Mahieu Dionysia, P., Martens, J. G., Niederle, J.,
**Schouten, R.M.**, & Duivesteijn, W. (2024) [Generating MNAR missingness in image data, with
additional evaluation ofMisGAN](pdfs/BNAIC_MisGAN_MNAR.pdf). Accepted for presentation at BNAIC/BeNeLearn 2024. [See here for the presentation](pdfs/BNAIC2024_MNAR.pdf). 

- **Schouten, R.M.**, Stevens, G.W.J.M., van Dorsselaer, S.A.F.M., Duinhof, E.L., Monshouwer, K., Pechenizkiy, M. & Duivesteijn, W. (2024) [Analyzing the interplay between societal trends and socio-demographic variables with local pattern mining: Discovering exceptional trends in adolescent alcohol use in the Netherlands](pdfs/Alcohol_BNAIC_2024.pdf). Accepted for presentation at BNAIC/BeNeLearn 2024. [See here for the presentation](pdfs/BNAIC2024_Alcohol.pdf).

- **Schouten, R.M.**, Duivesteijn, W., Rasanen, P, Paul, J.M., & Pechenizkiy, M. (2024) [Exceptional Subitizing Patterns: Exploring Mathematical Abilities of Finnish Primary School Children with Piecewise Linear Regression](pdfs/FUNA_ECMLPKDD_2024.pdf). In: Proceedings of the European Conference on Machine Learning and Principles and Practice of Knowledge Discovery in Databases (ECML PKDD), pp. 66-82.  

- **Schouten, R.M.**, Tascau, V., Ziegler, G.G., Casano, D., Ardizonne, M., & Erotokritou M.A. (2023) [Dropping incomplete records is (not so) straightforward](pdfs/2023_Schouten_MD.pdf). In: Proceedings of the 21st International Symposium on Intelligent Data Analysis (IDA), pp. 379-391.

- Verhaegh, R.F.A., Kiezebrink, J.J.E., Nusteling, F., Rio, A.W.A, Bendicsek, M.B., Duivesteijn, W. & **Schouten, R.M.** (2022) [A Clustering-inspired Quality Measure for Exceptional Preferences Mining — Design Choices and Consequences](pdfs/2022_Verhaegh_EPM.pdf). In: Proceedings of the International Conference on Discovery Science (DS), pp. 429–444.

- **Schouten, R.M.**, Zamanzadeh, D. & Singh, P. (2022) [Pyampute: a Python library for data amputation](https://rianneschouten.github.io/pyampute/build/html/index.html). Zenodo.  [https://doi.org/10.25080/majora-212e5952-03e](https://doi.org/10.25080/majora-212e5952-03e).

- Van der Haar, J.F., Nagelkerken, S.C., Smit, I.G., van Straaten, K., Tack, J.A., **Schouten, R.M.** & Duivesteijn, W. (2022) [Efficient Subgroup Discovery Through Auto-Encoding](pdfs/2022_vdHaar_AE.pdf) In: Proceedings of the 20th International Symposium on Intelligent Data Analysis (IDA), pp. 327-340.

- **Schouten, R.M.**, Duivesteijn, W. & Pechenizkiy, M. (2022) [Exceptional Model Mining for Repeated Cross-Sectional Data (EMM-RCS)](pdfs/2022_Schouten_EMMRCS.pdf). In: Proceedings of SIAM International Conference on Data Mining (SDM), pp. 585-593.

- **Schouten, R.M.**, Bueno, M.L.P., Duivesteijn, W. & Pechenizkiy, M. (2022) [Mining Sequences with Exceptional Transition Behaviour of Varying Order using Quality Measures based on Information-Theoretic Scoring Functions](pdfs/MiningExceptionalSequences_DAMI_2022.pdf) Data Mining and Knowledge Discovery (DAMI), 36: 379-413.

- IJsselhof R, Duchateau S, **Schouten R.M.**, Slieker M, Hazekamp M & Schoof P. (2020) [Long-Term Follow-Up of Pericardium for the Ventricular Component in Atrioventricular Septal Defect Repair](https://pubmed.ncbi.nlm.nih.gov/33164688/) World Journal for Pediatric and Congenital Heart Surgery, 11(6): 742-747.

- IJsselhof R.J., Duchateau S.D.R., **Schouten R.M.**, Freund, M.W., Heuser, J., Fejzic, Z., Haas, F., Schoof, P.H. & Slieker, M.G. (2019) [Follow-up After Biventricular Repair of the Hypoplastic Left Heart Complex](https://academic.oup.com/ejcts/article-abstract/57/4/644/5607276) European Journal of Cardiothoracic Surgery, 57(4): 644-651.

- **Schouten R.M.**, Lugtig, P. & Vink, G. (2018) [Generating missing values for simulation purposes: A multivariate amputation procedure](https://www.tandfonline.com/doi/full/10.1080/00949655.2018.1491577) Journal of Statistical Computation and Simulation, 88(15): 1909-1930.

- **Schouten, R.M.** and Vink, G. (2021) [The dance of the mechanisms: How observed information influences the validity of missingness assumptions](https://journals.sagepub.com/doi/full/10.1177/0049124118799376) Sociological Methods \& Research, 50(3): 1243-1258.

- Kappen, I.F.P.M., Bittermann, G.K.P., **Schouten, R.M.**, Bittermann, D., Etty, E., Koole, R., Kon, M., Van der Molen, M. & Breugem, C.C. (2017) [Long-term mid-facial growth of patients with a unilateral complete cleft of lip, alveolus and palate treated by two-stage palatoplasty: cephalometric analysis](https://www.researchgate.net/publication/308181181_Long-term_mid-facial_growth_of_patients_with_a_unilateral_complete_cleft_of_lip_alveolus_and_palate_treated_by_two-stage_palatoplasty_cephalometric_analysis) Clinical Oral Investigations, 21: 1801-1810.

- de Vries, C.P., **Schouten, R.M.**, Van der Kuur, J., Gottardi, L., & Akamatsu, H. (2016) [Microcalorimeter pulse analysis by means of principle component decomposition](https://arxiv.org/abs/1607.00834) In: Proceedings SPIE 9905, Space TElescopes and Instrumentation 2016: Ultraviolet to Gamma Ray, 99055v. DOI: 10.1117/12.2231627

---

### Awards and recognition

- Recognition as Excellent reviewer Research Track ECML PKDD 2024

- Award for Excellent course evaluation *Research Topics in Data Mining* 2022\/2023

- Award for Excellent course evaluation *Research Topics in Data Mining* 2021\/2022

---

### Grants and funding

|Year|Call|Type|Title|Together with|
|----|----------|-----------|---------------|------------------------|
|2024|Take-off Phase 1 NWO|€40k Feasibility study|Integration of Local Pattern Mining in Digital Assessment Tools|Prof. Mykola Pechenizkiy|
|2024|ECML PKDD Proceedings Chair|€500 reimbursement of entree ticket|Proceedings Chair|Dr. Wouter Duivesteijn|
|2022|AI for Health EWUU Alliance|EUR 45k seed money|Better Imputation by Generative Adversarial NeTworks (BIGANT)|Prof. Stef van Buuren, Dr. Gerko Vink, Hanne Oberman, Prof. Mykola Pechenizkiy, Prof. Cassio de Campos, **Rianne Schouten**,  Daniel Oberski, Dr. Thomas Debray, Prof. Fred van Eeuwijk|

--- 

### Domain collaborations

|Year|Organization|Type of collaboration|
|------|----------------------|------------------------|
|2025-now|AlgebraKit|Answering domain questions using EMM|
|2022-now|Turku Research Institute for Learning Analytics|Answering domain questions using EMM|
|2022-now|Dutch south west Psoriatic Arthritis Registry, Erasmus MC|Solving MD problems + Answering domain questions using EMM|
|2020-now|Biomedical Systems and Signals Group, Utwente + Hospital Twente|Using 
|2021-2024|Netherlands Institute of Mental Health and Addiction|Answering domain questions using EMM|
|2017-2019|University Medical Center Utrecht|Statistics consultant|

---

### Research visits
|Year|Where?|Who?|Why?|Duration|
|---|----|-----|-----------|---|
|2025|Ghent University|Prof. Tijl de Bie, Prof. Jefrey Lijfijt|To collaborate on Pattern mining and fairness|1 day|
|2025|Bielefeld University|Prof. Barbara Hammer|To collaborate on Concept drift, XAI and User interaction|1 week|
|2024|Radboud University Nijmegen|Dr. Marcos L.P. Bueno|To provide a guest-lecture in the AI for Healthcare course|1 day|
|2016|Columbia University|Prof. Andrew Gelman|To collaborate on Evaluating missing data methods, together with Dr. Gerko Vink|2 months|

---

### Teaching 

At Eindhoven University of Technology, my teaching track record is:

|Year|Course|Level|Activities|
|----|---------|-|----------------------|
|24\/25|Research Topics in Data Mining|MSc|Responsible for track: **Empirical Challenges in Data Mining**|
|22\/23|Research Topics in Data Mining|MSc|Taught 2 lectures about Missing Data, supervised 4 groups of students during research project|
|21\/22|Research Topics in Data Mining|MSc|Taught 2 lectures about Missing Data, supervised 7 groups of students during research project|
|20\/21|Foundations of Data Mining|MSc|Taught 2 lectures about Missing Data, provided answers in weekly Q\&A, developed practice questions for exam, administrative activities|

> “I took a lot of courses last year, but I like your instructions the most. It is not only because of your professional knowledge, but also because of your personality of being kind, patient, responsible.” (Jin Ouyang, Master student, 2022)

At Utrecht University, my teaching track record is:

|Year|Course|Level|Activities|
|----|------------|----|--------------|
|2016|Survey Research: Statistical Analysis and Estimation|Advanced MSc|Organization,Tutoring exercise classes|
|2016|Survey Research: Design, Implementation and Data Processing|Advanced MSc|Organization,Tutoring exercise classes|
|2015|Survey Research: Statistical Analysis and Estimation|Advanced MSc|Organization,Tutoring exercise classes|
|2015|Survey Research: Design, Implementation and Data Processing|Advanced MSc|Organization,Tutoring exercise classes|

> "Rianne was a first class assistant at our summer school courses. Not only was all material prepared extremely punctual and without errors, she also got very high student evaluations. I can wholeheartedly recommend Rianne!” (Prof. dr. Edith de Leeuw, 2016)

---

### Supervision 

Current students: 

|Start|Student|Type|Topic|
|----|----------|-------|-------------------|
|May 2025|Abdullahi Farah|MSc Thesis||
|Aug 2024|Lieke van den Biggelaar|PhD Research|Exceptional Model Mining with Time Series Data|
|Nov 2023|Emmanuel C. Chukwu|PhD Research|Counterfactual Explanations in Time Series Classification|

Alumni:

|Year|Student|Type|Topic|
|----|--------|-------|-------------------|
|24\/25|Haoqi Guo|MSc Thesis|Improving Diversity and Feasibility of Counterfactual Explanations, supervision together with prof.dr. Mykola Pechenizkiy|
|23\/24|Lieke van den Biggelaar|MSc Thesis with Catharina Hospital|Discovering subgroups of patients with exceptional Atrium Fibrillation based on ECGs, supervision together with dr. Wouter Duivesteijn|
|22\/23|Victoria Tascau|MSc Thesis with DEPAR\/Erasmus Medical Centrum|Handling Missing Values in Longitudinal Medical Data, supervision together with dr. Wouter Duivesteijn|
|21\/22|Mika van Loon|MSc Thesis|Bootstrap Hypothesis Tests for Evaluating Subgroup Descriptions in Exceptional Model Mining, supervision together with dr. Wouter Duivesteijn|
|21\/22|Varun Kamat|Internship at Signify|Recommendation Tool for Component Database|
|21\/22|Isabel van den Heuvel|Research Proposal|Equivalence Testing for Developing Fair Machine Learning Algorithms, supervision together with Hilde Weerts|
|21\/22|Mats Verbraak|Research Proposal|Handling Missing Data in the Prediction Domain using Multiple Imputation|
|20\/21|Bart van Dooren|MSc Thesis with Philips|Predicting Cardiovascular Risk with Objective Physical Activity Measurements, supervision  together with prof.dr. Mykola Pechenizkiy|

---

### Community service

I reviewed for top-level data mining conferences and statistical journals
(DAMI, ECML PKDD, EWAF, JRSSB, SiM, BimJ). I was recognized as an **excellent reviewer** at ECML PKDD 2024. I served as **Proceedings Chair** at ECML PKDD 2024.

---

### Software development

#### 1. `ampute` in R-package **mice**

```{r, eval = FALSE}
library(mice)
?ampute
```

R-function `ampute` is the implementation of a multivariate amputation procedure: a method for generating missing data in complete datasets. With `ampute`, it is straightforward to generate missing values in multiple variables, with different missing data proportions and varying underlying missingness mechanisms. Read [the article](pdfs/ArticleAmpute.pdf) or [the vignette](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html) to learn more.

#### 2. `parlMICE` 

For large datasets or when you want to impute with a large number of imputations, multiple imputation with `mice` in R-package **mice** may have a long run time. As a solution, Gerko Vink and I created wrapper function `parlMICE`, which allows for a parallel run of `mice`.

The function is now part of package **mice** under the name `parlmice`.

```{r, eval = FALSE}
library(mice)
?parlmice
```

All information can be found [in the github repo](https://github.com/RianneSchouten/parlMICE) or in the [vignette](https://www.gerkovink.com/parlMICE/Vignette_parlMICE.html).

#### 3. `pyampute`: the first Python library for data amputation

Library `pyampute` provides the multivariate amputation methodology to the Python community, and it does more. It has improved default settings, allows for a combined MAR+MNAR mechanism, for custom probability functions and since it is compatible with scikit-learn's `fit` and `transform` paradigm, seamless integration in data processing pipelines becomes easy. 

Find plenty of examples in [`pyampute`'s documentation](https://rianneschouten.github.io/pyampute/build/html/index.html). Davina's presentation at SciPy22 can be found [here](https://www.youtube.com/watch?v=iI0gx52ucMA&t=25560s).

Install using pip or from source:

```
pip install pyampute
```

```
git clone https://github.com/RianneSchouten/pyampute.git
pip install ./pyampute
```

---

### Invited presentations

|Year|Occasion|Type|Topic|Link to materials|
|----|------|------|------------|----|
|2025|Research visit @ Bielefeld University|Presentation|Exceptional Model Mining for Hierarchical Data||
|2025|Info-topic about my PhD research|News|Exceptional Model Mining for Hierarchical Data|[link](https://www.cursor.tue.nl/en/background/2025/januari/week-3/home-stretch-discovering-unusual-patterns-in-data/)|
|2024|AI for Healthcare|Invited guest-lecture at Radboud University Nijmegen|Exceptional Model Mining||
|2021|EAISI Eindhoven|Contribution to seminar|Towards a better understanding of exceptional lifestyle behaviour||
|2019|Workshop R-Ladies Amsterdam|Invited presentation at seminar|Developed and presented a workshop about analysis of missing values, evaluation and implementation of missing data methods|[link](https://github.com/RianneSchouten/workshops/tree/master/R-Ladies%20Workshop)|
|2018|ICT Open|Contribution to 1-day conference|Handling Missing Data in Data Science|[link](pdfs/ictopen2018_RianneSchouten.pdf)|
|2018|European Women in Technology|Masterclass at conference|Dealing with missing data in `R`: Amputation or Imputation?|[presentation and exercises](https://github.com/RianneSchouten/workshops/tree/master/WomenInTechnology)|
|2018|sat-R-Day|Contribution to 1-day conference|Missing data|[link](pdfs/SatRdayRianneSchouten.pdf)|
|2018|Data Science Hackathon|By invitation|Developed and lead a missing data challenge|[link](pdfs/Data_Science_Hackathon_Rianne_Schouten.pdf)|
|2017|Amst-R-Dam|Contribution to seminar|How to use R-function `ampute` to generate missing values in complete datasets|[article](pdfs/ArticleAmpute.pdf) and [documentation](https://rianneschouten.github.io/mice_ampute/vignette/ampute.html)|
|2017|UseR!2017|Contribution to conference|Introduction to multivariate amputation with `ampute`||
